# Load task data

## Study 2

The following participants are excluded from further analysis at this point, because of incomplete data:

* `S03`, `S14`, `S29`, `S38`, `S43`, `S46`: their T1 performance in session 1 was less than 63% correct, so they were not invited back. This cutoff was determined based on a separate pilot study with 10 participants. It is two standard deviations below the mean of that sample.
* `S25` has no data for session 2, as they stopped responding to requests for scheduling the session
* `S31` was excluded as a precaution after session 1, as they developed a severe headache and we could not rule out the possibility this was related to the tDCS

```{r load study 2 data}
dataDir_study2 <- here("data") # root folder with AB task data
subs_incomplete <- c("S03", "S14", "S25", "S29", "S31", "S38", "S43", "S46") # don't try to load data from these participants
df_study2 <- load_data_study2(dataDir_study2, subs_incomplete) %>%
  filter(complete.cases(.)) %>% # discard rows with data from incomplete subjects
   # recode first.session ("anodal" or "cathodal") to session.order ("anodal first", "cathodal first")
  mutate(first.session = parse_factor(paste(first.session, "first"), 
                                      levels = c("anodal first", "cathodal first"))) %>%
  rename(session.order = first.session)
n_study2 <- n_distinct(df_study2$subject) # number of subjects in study 2
```

```{r print slice study 2 data}
kable(head(df_study2,13), digits = 1, caption = "Data frame for study 2")
```

The data has the following columns:

* __subject__: Participant ID, e.g. `S01`, `S12`
* __session.order__: Whether participant received `anodal` or `cathodal` tDCS in the first session (`anodal_first` vs. `cathodal_first`).
* __stimulation__: Whether participant received `anodal` or `cathodal` tDCS
* __block__: Whether data is before (`pre), during (`tDCS`) or after (`post`) tDCS
* __lag__: Whether T2 followed T1 after two distractors (lag `3`) or after 7 distractors (lag `8`)
* __trials__: Number of trials per lag that the participant completed in this block
* __T1__: Proportion of trials (out of `trials`) in which T1 was identified correctly
* __T2__: Proportion of trials (out of `trials`) in which T2 was identified correctly
* __T2.given.T1__: Proportion of trials which T2 was identified correctly, out of the trials in which T1 was idenitified correctly (T2|T1).

The number of trials vary from person-to-person, as some completed more trials in a 20-minute block than others (because responses were self-paced):

```{r trial counts}
df_study2 %>%
  group_by(lag) %>%
  summarise_at(vars(trials), funs(mean, min, max, sd)) %>%
  kable(caption = "Descriptive statistics for trial counts per lag", digits = 0)
```


## Study 1

These data were used for statistical analysis in London & Slagter (2015)[^ref_london], and were processed by the lead author:

[^ref_london]: London, R. E., & Slagter, H. A. (2015). Effects of transcranial direct current stimulation over left dorsolateral pFC on the attentional blink depend on individual baseline performance. _Journal of Cognitive Neuroscience, 27(12)_, 2382-2393. doi: [10.1162/jocn_a_00867](https://doi.org/10.1162/jocn_a_00867)

```{r read study 1 data, output_max_height = "300px"}
dataPath_study1 <- here("data","AB-tDCS_study1.txt")
data_study1_fromDisk <- read.table(dataPath_study1, header = TRUE, dec = ",")
glimpse(data_study1_fromDisk)
```

We'll use only a subset of columns, with the header structure `block/stim`\_`target`\_`lag`\_`prime`, where:

* __block/stim__ is either:
    1. `vb`: "anodal" tDCS, "pre" block (before tDCS)
    2. `tb`: "anodal" tDCS, "tDCS" block (during tDCS)
    3. `nb`: "anodal" tDCS, "post" block (after tDCS)
    4. `vd`: "cathodal" tDCS, "pre" block (before tDCS)
    5. `td`: "cathodal" tDCS, "tDCS" block (during tDCS)
    6. `nd`: "cathodal" tDCS, "post" block (after tDCS)
* __target__ is either:
    1. `T1` (T1 accuracy): proportion of trials in which T1 was identified correctly
    2. `NB` (T2|T1 accuracy): proportion of trials in which T2 was identified correctly, given T1 was identified correctly
* __lag__ is either:
    1. `2` (lag 2), when T2 followed T1 after 1 distractor 
    2. `4` (lag 4), when T2 followed T1 after 3 distractors
    3. `10`, (lag 10), when T2 followed T1 after 9 distractors
* __prime__ is either:
    1. `P` (prime): when the stimulus at lag 2 (in lag 4 or lag 10 trials) had the same identity as T2 
    2. `NP` (no prime) when this was not the case. Study 2 had no primes, so we'll only keep these.
    
We'll also keep two more columns: `fileno` (participant ID) and `First_Session` (`1` meaning participants received anodal tDCS in the first session, `2` meaning participants received cathodal tDCS in the first session).

### Reformat

Now we'll reformat the data to match the data frame for study 2:

```{r Function to format data from study 1 as in study 2}
```

```{r format as for study 2}
df_study1 <- format_study2(data_study1_fromDisk)
n_study1 <- n_distinct(df_study1$subject) # number of subjects in study 1
```

```{r print slice study 1 data}
kable(head(df_study1,19), digits = 1, caption = "Data frame for study 1")
```
