---
title: AB_tDCS-EEG
author: Leon Reteig
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::html_clean:
    highlight: pygments
---

# Setup environment

```{r setup}

# Load packages
library(tidyverse)  # importing, transforming, and visualizing data frames
library(here) # (relative) file paths
library(knitr) # R notebook output
# Print version info
print(sessionInfo())
# Source functions
source(here("src", "func", "behavioral_analysis.R"))
```

# Load study 2 data

The following participants are excluded from further analysis at this point:

* `S03`, `S14`, `S29`, `S38`, `S43`, `S46`: their T1 performance in session 1 was less than 63% correct. This cutoff was determined based on a separate pilot study with 10 participants. It is two standard deviations below the mean of that sample.
* `S25` has no data for session 2, as they stopped responding to requests for scheduling the session
* `S31` was excluded as a precaution after session 1, as they developed a severe headache and we could not rule out the possibility this was related to the tDCS

```{r load study 2 data}
dataDir <- here("data","main") # root folder with AB task data
subs_incomplete <- c("S03", "S14", "S25", "S29", "S31", "S38", "S43", "S46") # don't try to load data from these participants
groupData <- load_data_study2(dataDir, subs_incomplete)
```

Preprocess

```{r print slice study 2 data}
kable(head(groupData,13), digits = 1, caption = "Data frame for study 2")
```

