---
title: AB-tDCS
subtitle: Correlation between anodal. vs. cathodal change scores 
author: Leon Reteig
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::html_clean:
    highlight: pygments
    gallery: true
    includes:
      in_header: header.html # prevent overlap with navbar
    
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup}

# Load packages
library(tidyverse)  # importing, transforming, and visualizing data frames
library(here) # (relative) file paths
library(knitr) # R notebook output
library(scales) # Percentage axis labels 
library(broom) # transform model outputs into data frames
library(ggm) # partial correlations
library(BayesFactor) # Bayesian statistics
library(knitrhooks) # printing in rmarkdown outputs
  output_max_height()
# Source functions
source(here("src", "func", "behavioral_analysis.R")) # loading data and calculating measures
load_data_path <- here("src","func","load_data.Rmd") # for rerunning and displaying
knitr::read_chunk(here("src", "func", "behavioral_analysis.R")) # display code from file in this notebook
```

```{r session info, output_max_height = "300px"}
print(sessionInfo())
```

```{r load data, child=load_data_path}
```

# Anodal vs. cathodal

Recreate data from the previous notebook

```{r recreate data}
# Recreate data
ABmagChange_study1 <- df_study1 %>%
  calc_ABmag() %>%# from the aggregated data, calculate AB magnitude
  calc_change_scores() # from the AB magnitude, calculate change scores
  
ABmagChange_study2 <- df_study2 %>%
  calc_ABmag() %>%# from the aggregated data, calculate AB magnitude
  calc_change_scores() # from the AB magnitude, calculate change scores
```

Study 1 reported that the "effects of" anodal and cathodal tDCS (`tDCS - baseline`) are anticorrelated: those participants that improve their performance (smaller AB magnitude) in the anodal session worsen in the cathodal session (and vice versa).

However, this correlation is not clearly present (and is not significant) in study 2:

## Plots {.tabset .tabset-fade .tabset-pills}

```{r Function to plot anodal vs cathodal change scores}

```

### Study 1

```{r anodal vs. cathodal plot study 1, fig.cap="Effect of anodal vs. effect of cathodal in study 1"}
plot_anodalVScathodal(ABmagChange_study1)
```

### Study 2

```{r anodal vs. cathodal plot study 2, fig.cap="Effect of anodal vs. effect of cathodal in study 2"}
plot_anodalVScathodal(ABmagChange_study2)
```

## Statistics {.tabset .tabset-fade .tabset-pills}

```{r Function to calculate partial correlations of anodal vs cathodal change scores}
```

### Study 1

Partial correlation between anodal and cathodal AB magnitude change score (tDCS - baseline), "controlling" for session order:

```{r partial correlation anodal vs cathodal study 1}
p_r_1 <- pcorr_anodal_cathodal(ABmagChange_study1)
r_study1 <- p_r_1$r
n_study1_pc <- n_study1 - 1 # lose one degree of freedom due to "controlling" for one variable in partial correlation
kable(p_r_1)
```

Zero-order correlation:

```{r zero-order correlation study 1}
r_1 <- tidy(cor.test(
  pull(filter(ABmagChange_study1, stimulation == "anodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  pull(filter(ABmagChange_study1, stimulation == "cathodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  method = "pearson"))
kable(r_1)

```

Bayes factor:

```{r Bayes factor study 1}
extractBF(correlationBF(
  pull(filter(ABmagChange_study1, stimulation == "anodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  pull(filter(ABmagChange_study1, stimulation == "cathodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score))) %>%
  select(bf) %>% # keep only the Bayes Factor
  kable()
```

### Study 2

Partial correlation between anodal and cathodal AB magnitude change score (tDCS - baseline), attempting to adjust for session order:

```{r partial correlation anodal vs cathodal study 2}
p_r_2 <- pcorr_anodal_cathodal(ABmagChange_study2)
r_study2 <- p_r_2$r
n_study2_pc <- n_study2 - 1 # lose one degree of freedom due to "controlling" for one variable in partial correlation
kable(p_r_2)
```

Zero-order correlation:

```{r zero-order correlation study 2}
r_2 <- tidy(cor.test(
  pull(filter(ABmagChange_study2, stimulation == "anodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  pull(filter(ABmagChange_study2, stimulation == "cathodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  method = "pearson"))
kable(r_2)
```

Bayes factor:

```{r Bayes factor study 2}
bf_study2 <- extractBF(correlationBF(
  pull(filter(ABmagChange_study2, stimulation == "anodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score),
  pull(filter(ABmagChange_study2, stimulation == "cathodal", measure == "AB.magnitude", change == "tDCS - baseline"), change.score))) %>%
  select(bf) # keep only the Bayes Factor

  kable(bf_study2)
```