\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\theequation}{S\arabic{table}} \setcounter{equation}{0}

```{r}
tDCS_AE <- read_csv2(here("data", "tDCS_AE.csv"), col_names = TRUE, progress = FALSE, col_types = cols(
  session = readr::col_factor(c("first", "second")), 
  stimulation = readr::col_factor(c("anodal", "cathodal"))))
```

```{r}
# Make long form data frame of sensation intensity
intensity <- tDCS_AE %>%
  select(everything(), -contains("conf"), -notes) %>% # drop other columns
  gather(sensation, intensity, itching:nausea) # make long form 

# Make long form data frame of sensation confidence
confidence <- tDCS_AE %>%
  select(contains("conf"), subject, session, stimulation) %>% 
  gather(sensation, confidence, conf.itching:conf.nausea) %>%
  # get rid of "conf." prefix so it matches the sensation intensity table
  mutate(sensation = str_replace(sensation, "conf.", "")) 

int_conf <- full_join(intensity,confidence)  %>%
  # gather confidence/intensity to make one plot for each
  gather(measure, rating, intensity, confidence) %>% 
  mutate(measure = factor(measure, levels = c("intensity", "confidence")))
```


```{r tab-tDCS-AE, results='asis'}
tbl_int_anodal <- intensity %>%
  filter(stimulation == "anodal") %>%
  count(sensation, intensity) %>%
  spread(intensity, n) %>%
  replace(is.na(.), 0) %>%
  mutate_if(is.double, as.integer)

tbl_int_cathodal <- intensity %>%
  filter(stimulation == "cathodal") %>%
  count(sensation, intensity) %>%
  spread(intensity, n) %>%
  replace(is.na(.), 0) %>%
  mutate_if(is.double, as.integer)

tbl_conf_anodal <- confidence %>%
  filter(stimulation == "anodal") %>%
  count(sensation, confidence) %>%
  spread(confidence, n) %>%
  replace(is.na(.), 0) %>%
  mutate_if(is.double, as.integer)

tbl_conf_cathodal <- confidence %>%
  filter(stimulation == "cathodal") %>%
  count(sensation, confidence) %>%
  spread(confidence, n) %>%
  replace(is.na(.), 0) %>%
  mutate_if(is.double, as.integer)

ae_colnames <- c(" ",
                 "none", "a little", "moderate", "strong", "very strong",
                 "n/a", "unlikely", "possibly", "likely", "very likely")
                 
bind_rows(full_join(tbl_int_anodal, tbl_conf_anodal, by  = "sensation"),
          full_join(tbl_int_cathodal, tbl_conf_cathodal, by  = "sensation")) %>%
  kable(col.names = ae_colnames, booktabs = TRUE,
        caption = "Number of reports of tDCS side effects") %>%
  kable_styling(bootstrap_options = "striped",
                latex_options =  "scale_down") %>%
  add_header_above(c(" " = 1, "Intensity rating" = 5, "Confidence rating" = 5)) %>%
  pack_rows(index = c("anodal session" = 8, "cathodal session" = 8)) %>%
  footnote(number = c(
    "Intensity rating: To which degree were the following sensations present during stimulation?", 
    "Confidence rating: To which degree do you believe this was caused by the stimulation?"))
```

```{r fig-tDCS-AE, fig.cap='(ref:caption-tDCS-AE)', fig.width=7.48031, fig.height=4.623, message=FALSE, warning=FALSE}
ae_plot <- ggplot(int_conf, aes(x = rating, fill = stimulation)) +
     facet_grid(measure ~ fct_reorder(sensation, rating, 
                                      .fun = function(x) sum(x > 0, na.rm = TRUE), .desc = TRUE)) +
    geom_bar(position = "stack") +
    stat_bin(binwidth = 1, geom = "text", size = geom_text_size, 
             aes(label = ..count..), position = position_stack(vjust = 0.25), check_overlap = TRUE) +
    scale_fill_manual(values = c("#F25F5C", "#4B93B1")) +
    xlim(0.5,4.5) + # exclude "0" ratings that were not present
    scale_y_continuous("number of sessions reported", limits = c(0,sum(!is.na(tDCS_AE$stimulation))),
                       # bound plot at max number of ratings
                       breaks = c(0,15,30,45,60,75,sum(!is.na(tDCS_AE$stimulation)))) +
  theme_minimal_hgrid(font_size = base_font_size, font_family = base_font_family) +
  theme(legend.background = element_blank(), legend.position = "top", 
        legend.direction = "horizontal", legend.title = element_blank()) +
  theme(strip.background = element_rect(fill = "grey90"), panel.spacing.y = unit(1,"lines"), 
        axis.ticks.x = element_line(), axis.ticks.length = unit(.25,"lines"))
suppressWarnings(print(ae_plot))
```

(ref:caption-tDCS-AE) __tDCS adverse events__. Number of reports out of 89 sessions (either anodal or cathodal tDCS). Top row shows intensity ratings [_little, moderate, strong, very strong_]; bottom row shows participant's confidence that event was related to tDCS [_unlikely, possibly, likely, very likely_]. Adverse events are sorted in descending order of number of reports (for very rare events (five reports or fewer for a given polarity), some text counts have been removed to prevent overlap).

| Difference                         | Study 2                                                       | Study 1                                     | Motivation                                                                                            |
|------------------------------------|---------------------------------------------------------------|---------------------------------------------|-------------------------------------------------------------------------------------------------------|
| Sample: size                       | 40                                                            | 34                                          |                                                                                                       |
| Sample: age                        | Mean: 20.94, SD: 4.25                                         | Mean: 22.4, SD: 2.8                         |                                                                                                       |
| Sample: gender                     | 29 female (73%)                                               | 20 female (59%)                             |                                                                                                       |
| Design: concurrent measurements    | EEG                                                           | None                                        |                                                                                                       |
| Design: inter-session interval     | 1 week                                                        | min. 48 hours                               | Increase similarity between sessions; longer washout time                                             |
| Design: trials per 20-minute block | Short-lag trials:   M=130, SD=17; Long-lag trials: M=65; SD=9 | Short-lag trials:40; Long-lag trials: 40    | Allow as many trials as possible in 20 minutes by making the task self-paced                          |
| Task: RSVP stream                  | 15 letters                                                    | 17 letters                                  | Shorten the total length of a trial, to allow completion of more trials                               |
| Task: inter-stimulus interval      | 1750 ms, + 1000 ms                                            | 480 ms                                      | Allow EEG responses to return to baseline                                                             |
| Task: priming condition            | Absent                                                        | Present                                     | Remove priming condition as it is orthogonal to main contrast, and showed no effects in @London2021   |
| Task: attentional blink conditions | lag 3, 8                                                      | lag 2, 4, 10                                | Increase proportion of trials where T2 was seen, for EEG analysis                                     |
| tDCS: conductive medium            | Conductive paste                                              | Saline solution                             | Prevent bridging EEG electrodes with (spreading) saline solution                                      |

Table: (\#tab:tab-diffs) Differences in the methodology and participant samples @London2021 (Study 1) and the present study (Study 2).